---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
```{r, results='hide', message=FALSE, warning=FALSE}
library(dplyr)
```

## Loading and preprocessing the data
The code bellow expects activity.zip to be extracted.

```{r, echo=TRUE}
data = read.csv('activity.csv')
data$date = as.Date(data$date)
head(data, n=5)
```

Five first rows does not tell us much about data set. Let's see random sample:

```{r, echo=TRUE}
data[sample(nrow(data), 5),]
```

## What is mean total number of steps taken per day?

```{r, echo=TRUE}
total_steps = sapply(split(data$steps,data$date), sum, na.rm=TRUE)
```

The mean and median of the total number of steps taken per day:

```{r, echo=TRUE}
mean(total_steps)
median(total_steps)
```

Total steps distribution:

```{r, echo=TRUE}
hist(total_steps, breaks=10)
```

## What is the average daily activity pattern?

```{r, echo=TRUE}
timeline = aggregate(steps ~ interval, data, mean)
plot(timeline, type='l', xlab="Interval", ylab="Steps")
title(main = 'Steps Taken over a Day')
```

```{r, echo=TRUE}
timeline = aggregate(steps ~ interval, data, mean)
print(timeline$interval[max(timeline$steps) == timeline$steps])
```

## Imputing missing values

```{r, echo=TRUE}
# Incomplete readings:
sum(!complete.cases(data))
# Complete readings:
sum(complete.cases(data))
```

Let's impute the missing data with the mean for that 5-minute interval: 

```{r, echo=TRUE, warning=FALSE}
avg_steps <- sapply(split(data$steps, data$interval), mean, na.rm=TRUE)
imputed <- ifelse(is.na(data$steps), avg_steps[as.character(data$interval)], data$steps)

imputed_data = mutate(data, steps = imputed)
```

```{r, echo=TRUE}
total_steps_imputed = sapply(split(imputed_data$steps,imputed_data$date), sum, na.rm=TRUE)
```

Total steps distribution:

```{r, echo=TRUE}
hist(total_steps_imputed, breaks=10)
```

The mean and median of the total number of steps taken per day:

```{r, echo=TRUE}
mean(total_steps_imputed)
median(total_steps_imputed)
```

**Both mean and median are higher. Inputing values increases the estimates of the total daily number of steps.**

## Are there differences in activity patterns between weekdays and weekends?

There is a difference. It seems that people walk more over the weekend.

```{r, echo=TRUE}
data$dow <- as.factor(ifelse(as.POSIXlt(data$date)$wday %in% c(0,6), 'weekend', 'weekday'))

library(lattice) 
aggregated = aggregate(steps ~ interval * dow, data, mean)
xyplot(steps ~ interval | dow, main="", ylab="Steps", xlab="Interval", data=aggregated, type="l", layout=c(1,2))
```

